---
title: "Occupational Prestige: A Reproducible Report"
author: "Your Name"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true 
    theme: cosmo
    code-fold: true
    self-contained: true
    output-file: index.html
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

if (!requireNamespace("pacman", quietly = TRUE))
  install.packages("pacman", repos = "https://cloud.r-project.org")

pacman::p_load(
  here,
  car,
  dplyr,
  ggplot2,
  vtable,
  kableExtra,
  broom,
  texreg,
  scales
)
```

```{r}
#| label: prep-data
#| include: false

# Load raw data
data(Prestige)
data_orig <- Prestige

# Recode variables
dfstudy1 <- data_orig %>%
  mutate(
    prestige_cont  = prestige,
    income_cont    = income,
    education_cont = education,
    women_pct      = women,
    income_cat = car::recode(income,
      "1000:8000 = 'Low'; 8001:12000 = 'Mid'; else = 'High'"),
    education_level = car::recode(education,
      "lo:10.09 = 'Low'; 10.1:13 = 'Mid'; 13.1:hi = 'High'"),
    prestige_tier = car::recode(prestige,
      "0:30.5 = 'Low'; 30.6:60 = 'Mid'; 60.1:100 = 'High'"),
    blue_collar  = car::recode(type, "'bc'   = 1; else = 0"),
    professional = car::recode(type, "'prof' = 1; else = 0"),
    high_women   = car::recode(women, "50:100 = 1; else = 0")
  ) %>%
  as.data.frame()

# Save processed data
dir.create(here("input/data/proc"), recursive = TRUE, showWarnings = FALSE)
dir.create(here("output/images"),   recursive = TRUE, showWarnings = FALSE)
dir.create(here("output/tables"),   recursive = TRUE, showWarnings = FALSE)
save(dfstudy1, file = here("input/data/proc/dfstudy1.RData"))

# Prepare regression data frame
df_reg <- dfstudy1 %>%
  mutate(
    income_cat      = factor(income_cat,      levels = c("Low", "Mid", "High")),
    education_level = factor(education_level, levels = c("Low", "Mid", "High")),
    high_women      = factor(high_women,  levels = c(0, 1), labels = c("No", "Yes")),
    blue_collar     = factor(blue_collar, levels = c(0, 1), labels = c("No", "Yes"))
  )

# Fit regression models
model_1 <- lm(prestige_cont ~ income_cont + education_cont, data = df_reg)
model_2 <- lm(prestige_cont ~ income_cont + education_cont + women_pct + blue_collar, data = df_reg)
model_3 <- lm(prestige_cont ~ income_cat + education_level, data = df_reg)
model_4 <- lm(prestige_cont ~ income_cat + education_level + high_women + blue_collar, data = df_reg)
```

## Introduction

This report analyses occupational prestige scores from the `Prestige` dataset (Duncan, 1961; `car` package).
The dataset covers **`r nrow(dfstudy1)`** occupations measured across Canada.
The mean prestige score is **`r round(mean(dfstudy1$prestige_cont), 1)`**
(SD = `r round(sd(dfstudy1$prestige_cont), 1)`),
with values ranging from `r min(dfstudy1$prestige_cont)` to `r max(dfstudy1$prestige_cont)`.

We examine how income, education, gender composition, and occupational type predict prestige.

## Data and Variables

### Descriptive Statistics

@tbl-descriptives reports summary statistics for all processed variables.

```{r}
#| label: tbl-descriptives
#| tbl-cap: "Descriptive statistics of processed variables"
#| echo: false

tab001 <- vtable::st(
  dfstudy1,
  labels         = TRUE,
  out            = "kable",
  summ           = c("notNA(x)", "mean(x)", "median(x)", "sd(x)", "min(x)", "max(x)"),
  summ.names     = c("N", "Mean (%)", "Median", "SD", "Min", "Max"),
  factor.numeric = FALSE,
  digits         = 2,
  title          = ""
) %>%
  kable_classic(html_font = "sans", full_width = FALSE) %>%
  footnote(general = "Numeric variables: mean, median, SD, min, max. Categorical: counts and percentages.")

kableExtra::save_kable(tab001, here("output/tables/table001.html"))
tab001
```

## Visualizations

### Univariate Distributions

@fig-hist shows the distribution of prestige scores across occupations.
The distribution is roughly symmetric with a slight right skew.
@fig-bar displays the frequency of income categories.

```{r}
#| label: fig-hist
#| fig-cap: "Distribution of occupational prestige scores"
#| fig-width: 7
#| fig-height: 4
#| fig-align: center
#| out-width: "85%"
#| dpi: 300

p_hist <- ggplot(dfstudy1, aes(x = prestige_cont)) +
  geom_histogram(bins = 15, fill = "steelblue", alpha = 0.7) +
  labs(x = "Prestige Score", y = "Frequency") +
  theme_minimal()

ggsave(here("output/images/figure001.png"), p_hist, width = 7, height = 4, dpi = 300)
p_hist
```

```{r}
#| label: fig-bar
#| fig-cap: "Frequency of income categories"
#| fig-width: 7
#| fig-height: 4
#| fig-align: center
#| out-width: "85%"
#| dpi: 300

p_bar <- ggplot(dfstudy1, aes(x = income_cat, fill = income_cat)) +
  geom_bar(show.legend = FALSE) +
  labs(x = "Income Category", y = "Count") +
  theme_minimal()

ggsave(here("output/images/figure002.png"), p_bar, width = 7, height = 4, dpi = 300)
p_bar
```

### Bivariate Relationships

@fig-scatter-edu presents the relationship between education and prestige, and
@fig-scatter-box shows the distribution of prestige across income categories.
Both panels suggest that higher education and income are associated with greater occupational prestige.

```{r}
#| label: fig-scatter-edu
#| fig-cap: "Prestige vs. years of education"
#| fig-width: 7
#| fig-height: 4
#| fig-align: center
#| out-width: "85%"
#| dpi: 300

p_scatter <- ggplot(dfstudy1, aes(x = education_cont, y = prestige_cont)) +
  geom_point(alpha = 0.6, size = 3, color = "steelblue") +
  geom_smooth(method = "lm", se = TRUE, color = "red") +
  labs(x = "Years of Education", y = "Prestige Score") +
  theme_minimal()

ggsave(here("output/images/figure003.png"), p_scatter, width = 7, height = 5, dpi = 300)
p_scatter
```

```{r}
#| label: fig-scatter-box
#| fig-cap: "Prestige score by income category"
#| fig-width: 7
#| fig-height: 4
#| fig-align: center
#| out-width: "85%"
#| dpi: 300

p_box <- ggplot(dfstudy1, aes(x = income_cat, y = prestige_cont, fill = income_cat)) +
  geom_boxplot(alpha = 0.7, show.legend = FALSE) +
  geom_jitter(width = 0.2, alpha = 0.3) +
  labs(x = "Income Category", y = "Prestige Score") +
  theme_minimal()

ggsave(here("output/images/figure004.png"), p_box, width = 7, height = 5, dpi = 300)
p_box
```

## Regression Analysis

We estimate four OLS models predicting occupational prestige.
Models 1 and 2 use continuous predictors; Models 3 and 4 use categorical versions.
The full model (Model 4) explains
**`r scales::percent(summary(model_4)$r.squared, accuracy = 0.1)`** of the variance in prestige
and includes gender composition and blue-collar status as additional controls.

Results are reported in @tbl-regression.

```{r}
#| label: tbl-regression
#| tbl-cap: "OLS regression models predicting occupational prestige"
#| results: asis
#| echo: false

coef_map <- list(
  "(Intercept)"          = "Intercept",
  "income_cont"          = "Income (continuous)",
  "education_cont"       = "Education (continuous)",
  "income_catMid"        = "Income: Mid (ref. Low)",
  "income_catHigh"       = "Income: High (ref. Low)",
  "education_levelMid"   = "Education: Mid (ref. Low)",
  "education_levelHigh"  = "Education: High (ref. Low)",
  "women_pct"            = "Women (%)",
  "high_womenYes"        = "High women share: Yes (ref. No)",
  "blue_collarYes"       = "Blue collar: Yes (ref. No)"
)

# Save standalone HTML file to output/tables/
htmlreg(
  list(model_1, model_2, model_3, model_4),
  custom.coef.map    = coef_map,
  custom.model.names = c("M1 Cont.", "M2 Cont.", "M3 Cat.", "M4 Cat."),
  caption            = "OLS regression models predicting occupational prestige",
  caption.above      = TRUE,
  file               = here("output/tables/table002.html")
)

# Render inline (fragment only — no file argument)
htmlreg(
  list(model_1, model_2, model_3, model_4),
  custom.coef.map    = coef_map,
  custom.model.names = c("M1 Cont.", "M2 Cont.", "M3 Cat.", "M4 Cat."),
  caption.above      = TRUE
)
```

## Conclusion

Using the `Prestige` dataset (`r nrow(dfstudy1)` occupations), we find that both income and education
are strong positive predictors of occupational prestige.
The best-fitting model (M4) achieves an adjusted R² of
`r round(summary(model_4)$adj.r.squared, 3)`,
confirming that categorical indicators alongside gender composition and occupational type
explain most of the variance in prestige scores.
